Cartogram Mapping and its Application to Cancer Data Visualization

### 2019-09-30 review_article.Rmd
hopefully fixed references

@@

<!-- interaction -->

@@

We would like to acknowledge Dr Earl Duncan (Research Associate at ARC Centre of Excellence for Mathematical & Statistical Frontiers, QUT). His time and effort given to edit and comment on this paper was invaluable.

We thank Dr Susanna Cramb (Spatial Modeller, Cancer Council Queensland) and Dr Peter Baade (Senior Research Fellow, Cancer Council Queensland) for providing the opportunity to cover this exploration of disease mapping methods in writing.
This literature review would not be possible without the opportunity provided by Queensland University of Technology, and Cancer Council Queensland, and the roles of Professor Kerrie Mengersen (Professor of Statistics, Science and Engineering Faculty, QUT) and Dr Earl Duncan in contributing to the Australian Cancer Atlas and supervision of Stephanie's Research Masters.

It was in the development of this online cancer atlas that methods for disease map displays, and visual communication strategies were explored. We are thankful for the opportunity to write about these visualizations and the situations in which they are appropriate.

Several R [@R] packages were used to produce this paper.
For data analysis the tidyverse package [@tidyverse] provided many useful functions.
ggplot2 package [@ggplot2] was used to create maps, with RColorBrewer package [@RColorBrewer] providing additional color palettes and ggthemes [@ggthemes] package providing themes.
The png package [@png] was used to access png images taken from online web atlases, cowplot package [@cowplot] was used to arrange these plots into grouped images.

The following packages were used to create transformations from the sf [@sf] geographical shapes of the states of America [@spData] and the Australian Statistical Areas (Level 3):
The cartogram [@cartogram] package creates contiguous, non-contiguous and Dorling cartograms.
Hexagon tile map displays were created using the sugarbag package [@sugarbag].

>>>>

@CPISACA provide weather maps as a thoroughly developed example of animation of spatial displays to communicate information to the general public. The movement of a weather system will follow a forecasted path, all map users can follow the animated path of the weather system across the geography over a specified period of time.

The Australian Cancer Atlas [-@TACA] provides [tours](
https://atlas.cancer.org.au/app/tour/lungcancer) which change the display to draw user's attention to areas on the map that are relevant to the story.
This implementation of animation gives users tools to plan their own exploration.

@@

<!-- interaction -->
Interaction with maps is an important component to public atlases. A purpose is to provide access to more information than is possible to display in a single map, without overwhelming the viewer. Too many choices though can similarly overwhelm a viewer, and thus decisions do need to be made about content to provide, or hide, for accurate and comprehensive communication of information. Similarly, providing ways for users to interact with the display encourages engagement, and creative, efficient, elegant, interactive tools elicits curiosity about the data.

@@

The authors would like thank Dr Earl Duncan, Professor Kerrie Mengersen, Dr Susanna Cramb and Dr Peter Baade for conversations on the content of this article.

The following R [@R] packages were used to produce this paper: tidyverse [@tidyverse], RColorBrewer [@RColorBrewer], ggthemes [@ggthemes], png [@png], cowplot [@cowplot], sf [@sf], spData [@spData], cartogram [@cartogram], sugarbag [@sugarbag], knitr [@knitr], rmarkdown [@rmarkdown].

Files to reproduce the paper, and code to reproduce the plots, are available at https://github.com/srkobakian/review.
